---
title: "R Notebook"
author: "Razmig Zeitounian"
output:
  html_document:
    df_print: paged
---

### Pokemon Data
```{r, message=FALSE}
#link: https://www.kaggle.com/abcsds/pokemon
pokemon <- read.csv("Pokemon.csv")

library(ggplot2)
library(car) #gives LeveneTest
library(EnvStats) #for Box-Cox transformations
library(MASS) # for model selecting
library(plotly)
library(knitr)
library(forcats)
library(rcompanion)
library(moments)

#we must convert Generation into a factor
pokemon$Generation <- as.factor(pokemon$Generation)
str(pokemon)
#notice that mega pokemon are counted as their own (e.g., AbomasnowMega)

kable(head(pokemon,15))
```

### Tables of Types
```{r}
#we make tables to get an idea of the amount of pokemon by each type
type.table <- table(pokemon$Type.1, pokemon$Type.2)
kable(type.table)

t1.table <- table(pokemon$Type.1)
kable(t1.table)

t2.table <- table(pokemon$Type.2)
kable(t2.table)
```
Above, we have a table showing the number of pokemon with an interaction between Type 1 and Type 2, type 1, and type 2.

## Plots and Summary Stats
```{r}
#see if generations have different numbers of Type 1's and base stat totals
ggplot(pokemon, aes(Type.1, Total, color = Generation)) + geom_boxplot() + theme(plot.subtitle = element_text(vjust = 1), 
    plot.caption = element_text(vjust = 1), 
    panel.grid.major = element_line(linetype = "blank"), 
    panel.background = element_rect(fill = "thistle1"), 
    plot.background = element_rect(fill = "cyan2")) +labs(title = "Stat Total by Type 1 and Generation", 
    x = "Type 1")

#now type 2:
ggplot(pokemon, aes(Type.2, Total, color = Generation)) + geom_boxplot() + theme(plot.subtitle = element_text(vjust = 1), 
    plot.caption = element_text(vjust = 1), 
    panel.grid.major = element_line(linetype = "blank"), 
    panel.background = element_rect(fill = "rosybrown1"), 
    plot.background = element_rect(fill = "aquamarine1")) +labs(title = "Stat Total by Type 2 and Generation", 
    x = "Type 2")

#view avg BST by generation in a plot
bst.by.gen <- aggregate(Total ~ Generation, data = pokemon, FUN = mean)
bst.by.gen.plot <- ggplot(bst.by.gen, aes(x=Generation, y=Total, group = 1)) + geom_point(col = "white") + geom_line(color="red")+ theme(plot.subtitle = element_text(vjust = 1), 
    plot.caption = element_text(vjust = 1), 
    panel.grid.major = element_line(linetype = "blank"), 
    panel.grid.minor = element_line(linetype = "blank"), 
    panel.background = element_rect(fill = "black"), 
    plot.background = element_rect(fill = "red")) + ylab("Average BST") + ggtitle("Average BST by Generation")

bst.by.gen.plot

#interactive version:
ggplotly(bst.by.gen.plot)

#
ggplot(pokemon, aes(Total, color=Generation, fill=Generation)) + geom_histogram() + theme(plot.subtitle = element_text(vjust = 1), 
    plot.caption = element_text(vjust = 1), 
    panel.background = element_rect(fill = "honeydew1"), 
    plot.background = element_rect(fill = "lightsteelblue1"))

### Plot frequency of each type
#fct_infreq orders from least to greatest for type I
ggplot(pokemon, aes(x=fct_infreq(Type.1))) + geom_bar(fill = "aquamarine", color = "red") + theme(plot.subtitle = element_text(vjust = 1), 
    plot.caption = element_text(vjust = 1), 
    panel.grid.minor = element_line(linetype = "blank"), 
    panel.background = element_rect(fill = "thistle1"), 
    plot.background = element_rect(fill = "orange1"))+labs(title = "Frequency of each Type I", 
    x = "Type I", y = "Count")

ggplot(pokemon, aes(x=fct_infreq(Type.2))) + geom_bar(fill = "aquamarine", color = "red") + theme(plot.subtitle = element_text(vjust = 1), 
    plot.caption = element_text(vjust = 1), 
    panel.grid.minor = element_line(linetype = "blank"), 
    panel.background = element_rect(fill = "ivory"), 
    plot.background = element_rect(fill = "tomato2"))+labs(title = "Frequency of each Type II", 
    x = "Type II", y = "Count")+labs(caption = "*Note that the empty label means no secondary type")
```
### Histograms for each stat by Generation
```{r}
#make histograms

#hp
hist.hp <- ggplot(pokemon, aes(x=HP)) + geom_histogram(col = "black", fill = "red") + facet_grid(.~Generation) + ggtitle("Distribution of HP stat by Generation")

#atk
hist.atk <- ggplot(pokemon, aes(x=Attack)) + geom_histogram(col = "red", fill = "blue") + facet_grid(.~Generation) + ggtitle("Distribution of Attack stat by Generation")

#def
hist.def <- ggplot(pokemon, aes(x=Defense)) + geom_histogram(col = "blue", fill = "aquamarine") + facet_grid(.~Generation) + ggtitle("Distribution of Defense stat by Generation")

#sp atk
hist.spatk <- ggplot(pokemon, aes(x=Sp..Atk)) + geom_histogram(col = "grey", fill = "yellow") + facet_grid(.~Generation) + ggtitle("Distribution of Special Attack stat by Generation")

#sp def
hist.spdef <- ggplot(pokemon, aes(x=Sp..Def)) + geom_histogram(col = "blue", fill = "royalblue2") + facet_grid(.~Generation) + ggtitle("Distribution of Special Defense stat by Generation")

#spd
hist.spd <- ggplot(pokemon, aes(x=Speed)) + geom_histogram(col = "yellow", fill = "blue") + facet_grid(.~Generation) + ggtitle("Distribution of Speed stat by Generation")

#######
ggplotly(hist.hp)
ggplotly(hist.atk)
ggplotly(hist.def)
ggplotly(hist.spatk)
ggplotly(hist.spdef)
ggplotly(hist.spd)
#notice how they are all approximately normal and may contain some outliets
```
### Base Stat Total (BST) Distribution
```{r}
hist.bst <- ggplot(pokemon, aes(x=Total)) + geom_histogram(col = "green", fill = "purple") + facet_grid(.~Generation) + ggtitle("Distribution of Base Stat Total (BST) by Generation") + theme(plot.subtitle = element_text(vjust = 1), 
    plot.caption = element_text(vjust = 1), 
    panel.background = element_rect(fill = "honeydew1"), 
    plot.background = element_rect(fill = "antiquewhite"))

ggplotly(hist.bst)
```
The histogram suggests BST is approximately normally distributed across all generations.

### One-Way ANOVA on BST using Generation
```{r}
pokemon.lm.model <- lm(Total ~ Generation, data = pokemon)
pokemon.aov <- aov(pokemon.lm.model)
summary(pokemon.aov)
#so no significant difference between at least one generation for BST
plot(pokemon.aov)
#our residuals look random, errors look normal, so this is good



#########################################
#One-Way ANOVA on Legendary and BST
legendary.model <- lm(Total ~ Legendary, data = pokemon)
legendary.aov <- aov(legendary.model)
summary(legendary.aov)

#This suggest legendaries have a higher BST than non-legendaries
```
### Visualize errors under the model
```{r}
ggplot(pokemon, aes(pokemon.lm.model$residuals)) + geom_histogram(color = "seagreen2") + xlab("Residuals")
```
This plot suggests our errors are skewed left, which violates the normality assumption of the residuals.

### Diagnostics/Formal Assumption Testing
```{r}
resid <- pokemon.lm.model$residuals
shapiro.wilk.test <- shapiro.test(resid)
shapiro.wilk.test
#so normality assumption is violated

brown.fors.test <- leveneTest(Total ~ Generation, data = pokemon, center = median)
brown.fors.test
#so equal variance by group condition is met
```

These results mean that although our equal variance condition is met for any reasonable value of $\alpha$, our errors are not approximately normally distributed. Therefore, we will consider transforming our variables to satisfy ANOVA assumptions.

### Transformations
```{r}
#check skew of response
skewness(pokemon$Total)

#try square rooting the Response Variable
transform.pok <- lm(sqrt(Total) ~ Generation, data = pokemon)
plot(transform.pok)

resid2 <- transform.pok$residuals
shapiro.wilk.test2 <- shapiro.test(resid2)
shapiro.wilk.test2
#so assumptions not met

#now try log(10)
transform.pok2 <- lm(log10(Total) ~ Generation, data = pokemon)
plot(transform.pok2)

resid3 <- transform.pok2$residuals
shapiro.wilk.test3 <- shapiro.test(resid3)
shapiro.wilk.test3
#assumptions not met
```
Since transformations did not satisfy ANOVA assumptions, we will use non-parametric methods for our hypothesis test.

### Non-Parametric Testing
```{r}
#add a rank column
pokemon$Rank = rank(pokemon$Total, ties = "average")

#get F-stat
F.OBS = summary(lm(Total ~ Generation, data = pokemon))$fstatistic["value"]

permuted.data = pokemon #So we don't overwrite the original data
permuted.data$Generation = sample(permuted.data$Generation, nrow(permuted.data), replace = FALSE) #Permuting the groups
Fi = summary(lm(Total ~ Generation, data = permuted.data))$fstatistic["value"]

#we will permute the data 3000 times
R = 3000
many.perms = sapply(1:R,function(i){
  permuted.data = pokemon #So we don't overwrite the original data
  permuted.data$Generation = sample(permuted.data$Generation, nrow(permuted.data), replace = FALSE) #Permuting the groups
  Fi = summary(lm(Total ~ Generation, data = permuted.data))$fstatistic["value"]
  return(Fi)
})

hist(many.perms, main = "Distribution for Permuted F values", xlab = "Fi")
points(y = 0, x = F.OBS, pch = 17)
#we have a distribution that is skewed right

#get p-val:
mean(many.perms >= F.OBS)

#try w/ Kruskall-Wallis test to compare:

nt <- nrow(pokemon)
Ri <- aggregate(Rank ~ Generation, data = pokemon, mean)$Rank
SR.2 = var(pokemon$Rank)
ni = aggregate(Total ~ Generation, data = pokemon, length)$Total

KW.OBS = 1/SR.2*sum(ni*(Ri - (nt+1)/2)^2) #Note, this assumes you calculate ni and Ri above
R = 3000
many.perms.KW = sapply(1:R,function(i){
  permuted.data = pokemon #So we don't overwrite the original data
  permuted.data$Generation = sample(permuted.data$Generation, nrow(permuted.data), replace = FALSE) #Permuting the groups
  SR.2 = var(permuted.data$Rank)
  ni = aggregate(Rank ~ Generation, data = permuted.data,length)$Rank
  Ri = aggregate(Rank ~ Generation, data = permuted.data,mean)$Rank
  KW.i= 1/SR.2*sum(ni*(Ri - (nt+1)/2)^2) 
  return(KW.i)
})
#p-val for KW:
p.value = mean(many.perms.KW > KW.OBS)
p.value
```
Both of these non-parametric tests lead to the conclusion that we cannot conclude the distribution of total (BST) is different for at least one generation.
