---
title: "fifa19"
author: "Razmig Zeitounian"
date: "6/22/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We start by importing libraries and the data.
[Click Here for Data source](https://www.kaggle.com/karangadiya/fifa19)

```{r prep fifa19-1, message=FALSE, warning=FALSE}
library(tidyverse)
library(countrycode)
library(readr)

fifa19data <- read_csv("C:/Users/razmi/OneDrive/Desktop/KaggleExploration/fifa19data.csv")

#drop columns that we are not interested in
fifa19data <- fifa19data %>% select(-c(X1, ID, Flag, `Club Logo`, Photo, `Loaned From`, LS:GKReflexes, `Release Clause`))
glimpse(fifa19data)

#remove euro sign and M from value column
new.value <- fifa19data %>% 
  pull(Value) %>% 
  substring(.,2) %>% 
  gsub("[M]", "", .) %>% 
  as.double(.) %>% 
  tibble() %>% 
  mutate(Value2 = .*10000) %>% 
  select(Value2)

#similarly, remove euro sign & k from wage column
new.wage <- fifa19data %>% 
  pull(Wage) %>% 
  substring(.,2) %>% 
  gsub("[K]", "", .) %>% 
  as.double(.) %>% 
  tibble() %>% 
  mutate(Wage2 = .*1000) %>% 
  select(Wage2)

new.weight <- fifa19data %>% 
  pull(Weight) %>% 
  gsub("[lbs]", "", .) %>% 
  as.numeric(.) %>% 
  tibble() %>% 
  mutate(Weight2 = .*1) %>% 
  select(Weight2)

#create a function to split up feet and inches, and then convert them to inches
col.to.split <- fifa19data$Height
height.calc <- function(col.to.split){
   split.height <- strsplit(col.to.split, "'")[1]
   #need to treat each part as a number
   feet <- as.numeric(split.height[[1]][1])
   feet.to.inches <- as.numeric(feet*12)
   inches <- as.numeric(split.height[[1]][2])
   height.inches <- as.numeric(feet.to.inches + inches)
   return(height.inches)
   }

#run function to create a new column to add later; it gives height in inches
height.inches <- unlist(lapply(fifa19data$Height, FUN = height.calc)) %>% 
  tibble() %>% 
  rename("HeightInch" = ".") 

#note: (weight2 is in lbs units) and (wage2 & value2 are in Euros); update df w/ new columns
fifa19data <- cbind(height.inches, new.value, new.wage, new.weight, fifa19data) %>% 
  select(-c(Value, Wage, Weight, Height))

#https://stackoverflow.com/questions/7680959/convert-type-of-multiple-columns-of-a-dataframe-at-once

#function changes data types 
convert.magic <- function(obj,types){
    for (i in 1:length(obj)){
        FUN <- switch(types[i],character = as.character, 
                                   numeric = as.numeric, 
                                   factor = as.factor)
        obj[,i] <- FUN(obj[,i])
    }
    obj
}

out <- convert.magic(fifa19data,c('numeric','numeric','numeric',
                                  'numeric', 'character', 'factor',
                                  'factor', 'numeric', 'numeric',
                                  'factor', 'numeric', 'factor',
                                  'factor', 'factor', 'factor',
                                  'factor', 'factor', 'factor', 
                                  'factor', 'factor', 'character',
                                  'factor'))
```

We are interested in what factors can determine overall rating. Therefore, we will start off by making a histogram
of what will be our response variable in this project.

```{r Overall Distribution}
#get overall statistic's mean to use to as a label later
# fifa19data$Overall <- as.numeric(fifa19data$Overall)
overall.mean <- mean(fifa19data$Overall)

hist <- ggplot(fifa19data, aes(Overall, label = mean(Overall))) + 
  geom_histogram(col = "aquamarine", fill = "maroon2", binwidth = 5) + 
  ggtitle("Histogram of Overall Stat of players for FIFA19") +
  ylab("Count") + 
  theme_light()

#make a histogram with a line showing the mean and label it
hist +
  geom_vline(aes(xintercept=overall.mean), linetype = "dashed") +
  geom_text(aes(0, overall.mean, label = paste0("Overall mean: ", prettyNum(overall.mean)), vjust = -15.5, hjust = -.3))
```

As we can see above, the $Overall$ statistic is approximately normally distributed. We will investigate to see if region has an effect on this as well.

```{r Distribution by Continent}
#we can investigate how continent has an impact on abilities

#reference: https://bradleyboehmke.github.io/tutorials/histograms

continent <- countrycode(sourcevar = fifa19data$Nationality, origin = "country.name",destination = "region")
df <- cbind(continent, fifa19data) %>% 
  drop_na(continent)

#calculate mean by region to use in histogram 
mean.by.region <- df %>% 
  group_by(continent) %>% 
  summarize(mean_overall = mean(Overall)) 

ggplot(df, aes(Overall)) +
  geom_histogram(color = "cadetblue", fill = "cornsilk2", binwidth = 5) + 
  facet_grid(.~continent) + 
  ggtitle("Distribution of Overall Statistic, by Continent") + 
  ylab("Count") + 
  geom_vline(data=mean.by.region, aes(xintercept=mean_overall)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 

#now as boxplot
ggplot(df, aes(continent, Overall)) +
  geom_boxplot(fill = "yellow", col = "royalblue") + 
  theme(axis.text.x = element_text(angle = 30, vjust = 0.5, hjust=1)) + 
  ggtitle("Boxplot of Overall Statistic, by Continent") +
  xlab("Continent") 
#or alternate format:
ggplot(df, aes(continent, Overall)) +
  geom_boxplot(fill = "yellow", col = "royalblue") + 
  ggtitle("Boxplot of Overall Statistic, by Continent") +
  xlab("Continent") + 
  coord_flip()
```

We can see that South Asian players have the lowest average Overall stat and Latin American & Carribean players have the highest Overall stat. The variances seem roughly equal for all continents, with the exception of South Asia, which has the lowest variance. We also see that Europe and Latin America/Carribean have more outliers than other regions, which suggests the best and worst players reside in these regions.


